Species richness and biomass temporal trends reveals different assembly processes of food-web in natural fish assemblages 

### 2021-10-12 paper/manuscript.Rmd
Update introduction and results

---
title: "Biomass and species richness temporal trends drives the assembly and the disassembly of food-web in natural fish assemblages"
author: Alain Danet, Maud Mouchet, Loubna El Madouri, Elisa Thébault and Colin Fontaine (ordre non fixé)
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
  bookdown::word_document:
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage[table]{xcolor}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
number_sections: true
---

@@

- TODO:
  - [ ] RM Pielou of abundances
  - [ ] Interaction tables 
  - [ ] Interaction plots

@@

## community structure is dynamic and affect ecosystem functions

Biodiversity affects ecosystem properties such as total biomass, productivity
and stability [@tilman_biodiversity_2014; @loreau_biodiversity_2001]. However,
current biodiversity is at risk, numerous studies reported dramatic declines of
population and community biomass [@ceballos_biological_2017;
@hallmann_more_2017; @albano_native_2021], as well as decrease in species
richness [@barnosky_has_2011; @newbold_global_2015]. Whereas numerous
ecological studies have shown that the different aspects of community structure
are tightly linked [@thebault_food-web_2003; @dunne_network_2006;
@soininen_relationship_2012], the consequences of temporal variations of community
biomass and species richness on community structure are not well-known.

## Variations of community biomass and species richness are predicted to affect community structure

Changes of species richness and community biomass are indeed
predicted to affect community structure. The positive link between biodiversity
and ecosystem functioning in experimental studies predict that loss and
gain of species richness should respectively result in loss and gain of
community biomass [@tilman_biodiversity_2014; @pennekamp_biodiversity_2018]. But
this relation is not always found in natural settings [@grace_integrative_2016;
@li_relationship_2018; @maureaud_biodiversityecosystem_2019].

Variations of species richness can also result in variation of evenness, i.e.
equitability of species abundances or biomasses. Several studies shows that
higher evenness of abundances is linked to higher plant biomass in grasslands
and primary productivity in marine ecosystems [@soininen_relationship_2012].
Furthermore, studies reported that higher evenness plays a role of insurance in
face of perturbations [@wittebolle_initial_2009]. Indeed, most studies reported
positive relationship between species richness and evenness, but also neutral
and negative ones [@soininen_relationship_2012; @ma_species_2005;
@maureaud_biodiversityecosystem_2019].

## Variations of community biomass and species richness are predicted to affect food-web structure

The distribution and the number of ecological interactions, i.e. the structure
of ecological networks, also covaries with species richness and community
biomass. In food-webs, connectance, which measures the degree of generalism in a
network, has been theoretically shown to positively scale with community biomass
[@thebault_food-web_2003].  Connectance has been also shown to be negatively
related to species richness in empirical food-webs [@dunne_network_2006].
Trophic level has also been shown to be positively linked to total biomass in
theoretical models [@wang_biodiversity_2018].

The different dimensions of network and community structure are closely linked
but, we have no direct evidence on how they covary according to the
loss or gain of species richness and community biomass across time.

@@

The links between the different component of community structure often come from
spatial gradient comparison. The space for time substitution allows to use the
studies mentioned above to predict how changes in species richness or community
biomass will in turn affect community and network structure. The hypothesis that
space can surrogate space is often implicitly made in ecological studies. This
method began to be widely used in the early 20th century to infer the temporal
succession of vegetation from space variations [@pickett_space-for-time_1989].
It supposes a homology between variations in space and in time i.e. that spatial
sequences corresponds to temporal sequences. This method have raised criticisms
[@damgaard_critique_2019], with studies showing analogous patterns in space and
time [@blois_space_2013] or not [@johnson_testing_2008;
@adler_contrasting_2007]. Spatial turnover of species composition, dispersal
limitation, disturbance and stochastic processes have been hypothesised to be
responsible for the failure of space for time substitution
[@adler_contrasting_2007; @blois_space_2013].

In this line, how community biomass and species richness temporal trends are
linked to temporal changes in food-web structure is still an open question, as
well as if space for time substitution works to infer consequences of species
richness and community biomass on food-web structure. The increasing
availability of ecological temporal series makes possible to assess those
questions.

Using a standardized monitoring of fish freshwater communities on the period
1995-2018, we assessed the relationships among temporal trends of food-web
structure, trends of community biomass and trends in species richness. We
then assessed if spatial relationships display the same patterns as temporal
ones, to assess the validity of space for time substitution.

@@

We inferred food-web structure in two steps [@bonnaffe_comparison_2021;
@danet_species_nodate]: (1) we first computed the metaweb, describing the
potential trophic interactions between all species size classes, (2) then we
extracted each local food-webs by extracting species size classes and their
interactions according their presence in a given sampling event.

@@

The feeding interactions between fishes were based on the body size ratio
between predators and preys. As this, we defined the predation window of a
piscivorous trophic species as 3% to 45% of its midpoint
[@mittelbach_ontogeny_1998; @claessen_impact_2002; @hart_handbook_2008].  We set
a trohic interaction  A trophic link was defined between the piscivorous trophic
species and all the trophic species whose midpoint in body size range was
included in the predation window.  The fish-resource trophic links were set
between the trophic species and the resources nodes according to the food items
of their size class.  Finally, the feed interactions between resources were
defined according to the literature [@allan_stream_2007; @hart_handbook_2008].

@@

We characterized food-webs by its connectance, describing the degree of
generalism in a network [@dunne_network_2006; @delmas_analysing_2019]. It is
computed as the ratio between the actual number of links and the number of
potential links ($C = \frac{L}{S(S-1)}$, with $L$ being the number of links and
$S$ the number of trophic species). We also computed the average trophic level,
weighted by the biomass of the trophic species. The trophic level of a trophic
species is computed recursively as 1 added to the mean trophic level of the
preys that it consumes.

@@

To characterize the links among the temporal trends of network and community
structure and the temporal trends of community biomass and species richness, we
characterized the temporal trajectories of network and community structure at
each site. We used a method based on second order polynomial which allows to
assess both the direction and the monotony of temporal trends
[@rigal_method_2020]. To avoid complex variations of biomass and species
richness, we selected the stations that displayed monotonic trends for these
variables, i.e. the ones for which the second order polynomial term was non
significant. Over the `r nb_st` stations, the selection procedure resulted in 
`r nb_st_rich_bm` for community biomass and species richness analysis.

There was a clear relationship between the size of the communities and their
temporal dynamics. Communities that displayed low community biomass at the
beginning of their monitoring tend to display increase in biomass through time
and reversely. There same the same pattern for species richness, with
communities that displayed high species richness at the beginning of their
monitoring tend to display decrease in species richness through time and
reversely.

@@

With the aim to test the relationship among the temporal dynamics of network
structure from one side and species richness and community biomass from another
side, we modelled the dynamics the temporal trends of food-web structure ($dN$)
in function of the temporal trends of community biomass ($dB$) and species
richness ($dR$).

$dN = \alpha + \beta_1dB + \beta_2dR + \beta_3dBdR + \epsilon$

@@

In order to control for the uncertainties of the estimated temporal trends, each
temporal trend was weighted by the inverse of the product of the standard errors of the
estimated trends, i.e. slopes for the response and the explicative
variables.

@@

$N =  B + R + E_1 + E_2$

@@

loadd(rsq_table_bm_rich_mono_stable_trends, sp_rsq_table_bm_rich_mono_stable_trends)
```

@@

Overall, we found high heterogeneity in the explicative power of statistical
models linking temporal dynamics of food-web structure and the one of community
biomass and species richness spanning from $R_{adj}^2= `r rsq_piel_nind`$ for
pielou of individual number to $R_{adj}^2= `r rsq_tlvl`$ for average trophic
level (Table \@ref(tab:rsq)). The spatial models were more performant, with
$R_{cond}^2$ varing from $`r rsq_tlvl_sp`$ to $`r rsq_piel_bm_sp`$ (Table
\@ref(tab:rsq-sp)).

@@

### Community structure and community biomass

@@

tps_bm_sest <- tps_bm$sest
names(tps_bm_sest) <- tps_bm$rep
```
We found a significant positive relationship between the temporal trends of
community biomass from one side and species richness, average trophic level,
connectance from the other side (Fig. \@ref(fig:bef) and \@ref(fig:bmr), Table \@ref(tab:aov),
respectively  $r_\delta = `r tps_bm_sest["rich"]`$, $`r tps_bm_sest["tlvl"]`$, 
$`r tps_bm_sest["ct"]`$), meaning that they temporally increased when the community
biomass temporally increased and respectively that they temporally decreased
when the community biomass decreased in time.

We found a significant negative relationship between the temporal trends of
community biomass and that of the pielou of biomass 
( $r_\delta = `r tps_bm_sest["p_bm"]`$), meaning that the eveness of biomass
across species temporally increased when the community biomass temporally
decreased and respectively that the evenness of biomass temporally decreased
when the community biomass temporally increased (Fig. \@ref(fig:bmr),
Table \@ref(tab:aov)).

---
#We also found the same pattern between the temporal
#trends of community biomass and that of the pielou of individual number
#($r_\delta = `r tps_bm_sest["p_n"]`$). 
---

---
#Overall, we did not find effect of assembly/disassembly on the relationships
#between the temporal trends of community structure and the one of community
#biomass. We also found that the median community biomass had no effect on the
#relationship between the temporal trends of community structure and the one of community
#biomass.
---

### Community structure and species richness

@@

The temporal trends of community biomass and pielou of biomass, had a positive
relationship with that of species richness (Fig. \@ref(fig:bmr), Table
\@ref(tab:aov), $r_\delta = `r tps_r_sest["p_bm"]`$) while we did not find
significant relationships between the temporal trends of average trophic level,
connectance and the one of species richness (Fig.
\@ref(fig:bmr), Table \@ref(tab:aov)).

---
#We also found that the relationship between the temporal trends of pielou of
#individual and connectance and that of species richness was dependant of the
#overall species richness of the community (Fig. \@ref(fig:rich), Table
#\@ref(tab:aov)), meaning that the relationship between temporal trends of
#pielou of individual and that of species richness was more positive as the
#community contained more species (Fig. \@ref(fig:rich), Table
#\@ref(tab:reg)). While the relationship between temporal trends of pielou
#of individual and that of species richness was more negative as the community
#contained more species (Fig. \@ref(fig:rich), Table \@ref(tab:reg)).
---

## Spatial relationships

@@

### Community structure and community biomass

We found that species richness, connectance, average trophic level were
spatially positively related to community biomass (resp. 
$r_\delta = `r sp_b_sest["rich"]`$, $`r sp_b_sest["ct"]`$, 
$`r sp_b_sest["tlvl"]`$, Fig.
\@ref(fig:bmr), Table \@ref(tab:aov-sp)) while the pielou of biomass was
negatively related to community biomass ($r_\delta = `r sp_b_sest["p_bm"]`$).

### Community structure and species richness

Community biomass, the eveness of biomass were
positively related to species richness (resp. 
$r_\delta = `r sp_r_sest["bm"]`$, $`r sp_r_sest["p_bm"]`$,
$`r sp_r_sest["p_n"]`$) while average
trophic level and connectance were negatively related to
species richness (resp. $r_\delta = `r sp_r_sest["tlvl"]`$, $`r sp_r_sest["ct"]`$,
Fig. \@ref(fig:bmr), Table \@ref(tab:aov-sp)).

@@

@@

We found that the temporal increase of the average trophic level was positively
linked to increase of community biomass in line with a previous model suggesting
that the presence of higher predator can lead to higher community biomass
[@wang_biodiversity_2018]. Increase of average trophic level can be associated
to either gain of biomass in the highest trophic level nor losses of biomass in
the lowest trophic. In our case, such gain or losses of average trophic level is
probably associated to the variations of the presence of large fishes. The
finding that the evenness of biomass, measured by the pielou, decreased through
time as community biomass decreased and reversely, meaning that biomass
distribution across species become less even when community biomass decrease and
reversely. It further indicates that temporal loss of community biomass are
linked with the loss of the higher trophic level.

We also found in space that evenness of biomass was positively related to
community while eveness of abundances had no relation (Fig. SXX). It
constrasts with previous studies which found a negative relationship between
biomass and eveness of abundances in marine and plant communities
[@drobner_relation_1998; @maureaud_biodiversityecosystem_2019] or studies that
found more even plant communities in terms of abundances number display higher
community biomass [@wilsey_biodiversity_2000].

We further found that the temporal trends of connectance was positively
linked to that one of community biomass, meaning that connectance increases through
time in a given site when community biomass increases through time and reversely.
This is in line with the fact that the fishes located at higher trophic level
also display a larger number of links as the size of the predation window
increases with the size of the fishes. This result is also in accordance with  
theoretical work showing that higher food-web connectance result in higher
community biomass [@thebault_food-web_2003].

@@

To the difference with community biomass, the temporal trends of eveness of
biomass was positively related to the temporal trends of species
richness, meaning that as species richness temporally increase, the
distribution of biomass become more even, while temporal trends of eveness of
individual had no significant relationship with species richness. Those results
indicate that biomass trends are linked to few big individuals while richness
trends are linked to overall shift in dominance in communities.

Surprinsingly, we found that the temporal trends of connectance was not related to
the one of species richness as found spatially in most empirical food-web
[@dunne_network_2006] and also in the present study. This result suggests that
the structure of food-webs might be not very sensitive to variation of species
richness and that species loss and gain concerns species of low trophic level,
which in turn does affect connectance very little.

## Temporal dynamics confirms space for time substitution BEF

Spatially and temporally, the species richness was positively linked to that of
community biomass. Spatially, this result is in accordance with numerous studies was
already found in streams [@woods_testing_2020; @danet_species_nodate] but also
in grasslands, salt marches and forests [@hector_plant_1999;
@tilman_biodiversity_2014; @wu_relationship_2015; @li_relationship_2018; but see
@grace_integrative_2016]. Therefore, our results confirm that BEF prediction
works also in time. The positive relationship of species richness on total
biomass might be due to resource use complementarity [@carey_determining_2011],
which is held by niche differentiation [@loreau_biodiversity_1998;
@loreau_partitioning_2001] or to the higher probability of getting high biomass
species with higher species richness [@loreau_biodiversity_2001].

@@

In contrast with a meta-analysis [@soininen_relationship_2012], we found that
the eveness of individual number is positively related to the one of species
richness in space in stream fish assemblages but found no relationship in time.

This result has profound implications for conservation. As more and more
studies report community biomass and species richness trends [@hallmann_more_2017;
@ceballos_biological_2017], we documented that they might not be
equivalent in terms of consequences on the structure of the community. In
particular, the loss of the highest trophic level associated to the loss of
community biomass could of have dramatic consequences on ecosystem functioning
as highest predators are prone to stabilise communities
[@shanafelt_stability_2018; danet_species_nodate]. In contrasts, the loss of
species richness was poorly associated to modification of food-web structure,
suggesting that the species loss and gain.

@@

```{r}
fig_mat_cap <- paste0(
  "(A) Distribution of the fishing sites. (B-F) Temporal trends of biomass,
  species richness, connectance, average trophic level and eveness of biomass
  for two stations."
)
```
```{r, fig.cap=fig_mat_cap, out.width="100%"}
knitr::include_graphics(get_mypath("figures", "matmet.png"))
```

>>>>

---
title: "Species richness and biomass temporal trends reveals different assembly processes of food-web in natural fish assemblages " 
author: Alain Danet, Loubna El Madouri, Maud Mouchet, Elisa Thébault and Colin Fontaine (ordre non fixé)
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
  bookdown::word_document2:
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage[table]{xcolor}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
number_sections: true
---

@@

@@

## Trends of biodiversity needs to be assessed for multiple facets

Multiple studies reported dramatic declines of population and community biomass
[@ceballos_biological_2017; @hallmann_more_2017; @albano_native_2021], as well
as decrease in species richness [@barnosky_has_2011; @newbold_global_2015] and
turnover in species composition [@comte_climate_2021]. The documentation of
biodiversity temporal trends, is crucial to evaluate the effect of
global changes [but see @cardinale_is_2018]. Unfortunately, studies often
report trends of one dimension of biodiversity at a time [@hallmann_more_2017,
etc, but see @]. Because the functioning of ecosystem depends ultimately on
multiple facets of biodiversity, such as species richness, evenness and
functional composition [REF @thibaut_understanding_2013], the monitoring of one
biodiversity trend is not enough to project potential effects on ecosystem
functioning.

@@

Another concern is that the links between the different components of community
structure often come from spatial comparison [@dunne_network_2006;
@soininen_relationship_2012; @grace_integrative_2016 @danet_species_nodate]. In
order to use spatial comparison to predict the fate of communities facing
changes in species richness or community biomass, studies often implicitly use
the "space for time substitution" hypothesis. This method began to be widely
used in the early 20th century to infer the temporal succession of vegetation
from space variations [@pickett_space-for-time_1989]. It supposes a homology
between variations in space and in time i.e. that spatial sequences correspond
to temporal sequences.  This method have raised criticisms
[@damgaard_critique_2019], with studies showing analogous patterns in space and
time [@blois_space_2013] or not [@johnson_testing_2008;
@adler_contrasting_2007]. Spatial turnover of species composition, dispersal
limitation, disturbance and stochastic processes have been hypothesised to be
responsible for the failure of space for time substitution
[@adler_contrasting_2007; @blois_space_2013].

## Temporal variation of food-webs is largely unknown

The documentation of biodiversity trends generally neglect species interactions
although they are fundamentals to ecosystem functioning
[@thebault_food-web_2003; oconnor_general_2016]. In food-webs, empirical and
theoretical studies reported generally that trophic level positively scales to
community biomass [@wang_biodiversity_2018; @danet_species_nodate] (add
mechanisms ?). Trophic level has also been shown to stabilise the temporal
variation of community biomass [@shanafelt_stability_2018;
@danet_species_nodate]. Connectance, which measures the degree of generalism in
a network, was shown to increase robustness to extinction [@dunne_network_2002].
However, the temporal variation of food-web structure is largely unknown
[@kortsch_disentangling_2021]. Food-web temporal variation are often reported
across small time step [warren_spatial_1969; @pecuchet_species_2020;
@kaartinen_high_2012].

We know very little how food-web structure vary across time and particularly how
those variations are linked to the variations of other aspects of community
structure such as species richness and community biomass. Spatially, connectance
has been shown to be negatively related to species richness in empirical
food-webs [reviewed in @dunne_network_2006]. Connectance has been shown to scale
positively or negatively with community biomass in theoretical models
[@thebault_food-web_2003; @poisot_trophic_2013] (*Add mechanism?*). In order to
improve our understanding on the assembly mechanisms in food-webs, it is thus
crucial to assess how biomass and species richness variation are associated to
variations in food-web structure.

In this line, how the temporal changes of community biomasse are related to
temporal changes of species richness can still be question in natural settings,
especially in food-webs. Our knowledge on how the temporal changes of food-web
structure are related  to temporal changes in community biomass and species
richness is almost zero. Our knowledge on the relationships between food-web
structure comes often from comparison of food-web snapshots along spatial
gradient, raising the question of the validity of space for time substitution to
infer consequences of species richness and community biomass on food-web
structure. The increasing amount of availability ecological temporal series
makes possible to assess those questions. Using a standardized monitoring of
fish freshwater communities on the period 1995-2018 and a method of food-web
inference based on allometric relationships and food diet datas, we assessed the
relationships among temporal changes of food-web structure, trends of community
biomass and trends in species richness. We then assessed the validity of space
for time substitution by comparing if temporal relationships display the same
patterns than relationships obtained along spatial gradient.

---
## Variations of community biomass and species richness are predicted to affect community structure
#Changes of species richness and community biomass are indeed
#predicted to affect community structure. The positive link between biodiversity
#and ecosystem functioning in experimental studies predict that loss and
#gain of species richness should respectively result in loss and gain of
#community biomass [@tilman_biodiversity_2014; @pennekamp_biodiversity_2018]. But
#this relation is not always found in natural settings [@grace_integrative_2016;
#@li_relationship_2018; @maureaud_biodiversityecosystem_2019].
#
#Variations of species richness can also result in variation of evenness, i.e.
#equitability of species abundances or biomasses. Meta-analysis reported that
#higher evenness of abundances is linked to higher plant biomass in grasslands
#and primary productivity in marine ecosystems [@soininen_relationship_2012].
#Furthermore, studies reported that higher evenness plays a role of insurance in
#face of perturbations [@wittebolle_initial_2009]. Indeed, most studies reported
#positive relationship between species richness and evenness, but also neutral
#and negative ones [@soininen_relationship_2012; @ma_species_2005;
#@maureaud_biodiversityecosystem_2019].
## Variations of community biomass and species richness are predicted to affect food-web structure
---

@@

We inferred food-web structure in two steps [@bonnaffe_comparison_2021;
@danet_species_nodate]: (1) we first computed the metaweb, describing the
potential trophic interactions between all species size classes, (2) then we
extracted each local time snapshot of food-webs by extracting species size
classes and their interactions according their presence in a given sampling
event.

@@

The feeding interactions between fishes were based on the body size ratio
between predators and preys. As this, we defined the predation window of a
piscivorous trophic species as 3% to 45% of its midpoint
[@mittelbach_ontogeny_1998; @claessen_impact_2002; @hart_handbook_2008].  A
trophic link was defined between the piscivorous trophic species and all the
trophic species whose midpoint in body size range was included in the predation
window.  The fish-resource trophic links were set between the trophic species
and the resources nodes according to the food items of their size class.
Finally, the feed interactions between resources were defined according to the
literature [@allan_stream_2007; @hart_handbook_2008].

@@

We characterized food-webs by their connectance, describing the degree of
generalism in a network [@dunne_network_2006; @delmas_analysing_2019]. It was 
computed as the ratio between the actual number of links and the number of
potential links ($C = \frac{L}{S(S-1)}$, with $L$ being the number of links and
$S$ the number of trophic species). We also computed the average trophic level,
weighted by the biomass of the trophic species. The trophic level of a trophic
species was computed recursively as 1 added to the mean trophic level of the
preys that it consumes.

@@

To characterize the links among the temporal trends of network and community
structure and the temporal trends of community biomass and species richness, we
characterized the temporal trajectories of network and community structure at
each site. We used a method based on second order polynomial which allows to
assess both the direction and the monotony of temporal trends
[@rigal_method_2020]. To avoid the inclusion of many types of biomass and species
richness trends (i.e. accelerating, decelerating, etc...), we selected the
stations that displayed monotonic trends for these variables, i.e. the ones for
which the second order polynomial term was non significant. Over the `r nb_st`
stations, the selection procedure resulted in `r nb_st_rich_bm` for community
biomass and species richness analysis. However, the results were robust to the
inclusion of non monotonic trends (Fig. SXX).

There was a clear relationship between the size of the communities and their
temporal dynamics. Communities that displayed low community biomass at the
beginning of their monitoring tend to display an biomass increase through time
and reversely. We observed the same pattern for species richness, with
communities that displayed high species richness at the beginning of their
monitoring tend to display a decrease in species richness through time and
reversely.

@@

We modelled the dynamics the temporal trends of food-web structure ($dN$)
in function of the temporal trends of community biomass ($dB$) and species
richness ($dR$) in order to to test the relationship among the temporal dynamics of network
structure from one side and species richness and community biomass from another
side.

$dN = \alpha + \beta_1dB + \beta_2dR + \epsilon$

@@

In order to control for the uncertainties of the estimated temporal trends, each
temporal trend was weighted by the inverse of the product of the standard errors of the
estimated trends, i.e. slopes for the response and the explicative
variables. We also tested a statistical model taking in account the interaction
between the temporal trends of species richness and community biomass (i.e.
adding $\beta_3dBdR$), the
results are presented in supplementary material (Fig. SXX).

@@

$N = \alpha + \beta_1B + \beta_2R + \beta_3E_1 + \beta_4E_2 + \epsilon$

@@

loadd(rsq_table_bm_rich_mono_stable_trends, sp_rsq_table_bm_rich_mono_stable_trends)

loadd(rigal_classification)
```

```{r}
class_trends_bm_rich <- rigal_classification %>%
  filter(variable %in% c("log_bm_std", "log_rich_std")) %>%
  unnest(classif) %>%
  select(variable, station, direction, acceleration, shape_class) 
res_trends <- class_trends_bm_rich %>%
  filter(acceleration == "constant") %>%
  group_by(variable, direction) %>%
  summarise(n = n()) %>%
  group_by(variable) %>%
  mutate(total = sum(n), percent = n / sum(n) * 100 %>% round(., 2))
```

@@

Overall, we found high heterogeneity in the explicative power of statistical
models linking temporal dynamics of food-web structure and the one of community
biomass and species richness spanning from $R_{adj}^2= `r rsq_piel_nind`$ for
the evenness of abundances to $R_{adj}^2= `r rsq_tlvl`$ for average trophic
level (Table \@ref(tab:rsq)). The spatial models were more performant, with
$R_{cond}^2$ varing from $`r rsq_tlvl_sp`$ to $`r rsq_piel_bm_sp`$ (Table
\@ref(tab:rsq-sp)). Over the station in the dataset, we found that
`r res_trends[res_trends$variable == "log_bm_std", ]$total` stations displayed monotonic trends for biomass. Among them, 
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "stable", ]$percent`%
were classified stable,
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "decrease", ]$percent`%
decreasing,
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "increase", ]$percent`%
increasing. Regarding the species richness, we found that
`r res_trends[res_trends$variable == "log_rich_std", ]$total` stations displayed monotonic trends.
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "stable", ]$percent`%
were classified stable,  
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "decrease", ]$percent`%
as decreasing, and 
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "increasing", ]$percent`% increasing.

@@

@@

tps_bm_sest <- tps_bm$sest
names(tps_bm_sest) <- tps_bm$rep
```

@@

@@

We found a strong positive relationship between the temporal trends of community
biomass and the one of species richness ($r_\delta = `r tps_r_sest["p_bm"]`$,
Fig. \@ref(fig:bef), Table \@ref(tab:aov)), meaning that both temporal trends
had the same direction in average. At stations
where species richness decreased over time, community biomass also decreased
over time. On the other hand, at stations where species richness increased over
time, community biomass also increased over time. In spatial analysis, we also
found a positive relationship between median community biomass and median
species richness, but twice less strong than the temporal one
($r_\delta = `r sp_b_sest["rich"]`$).

The temporal trends of connectance was positively linked with the one of
community biomass ($r_\delta = `r tps_bm_sest["ct"]`$), meaning also that both
temporal trends had the same direction in average. That is, the stations where
food-web connectance decreased over time had also community biomass that
decreased over time and reversely. It means that communities that lose biomass
over time, also loses generalist predators and reversely that communities
gaining biomass over time also gain generalist predators. Spatially, we found
also a positive relationship between median connectance and median species
richness ($r_\delta = `r sp_b_sest["ct"]`$). However, we did not observed
significant relationships between the temporal trends of connectance and the one
of species richness (Fig. \@ref(fig:bmr)D) when we found spatially a stong
negative relationship between median connectance and median species richness
($r_\delta = `r sp_r_sest["ct"]`$, Fig. \@ref(fig:std)).

We found that the temporal trends of average trophic level had a strong positive
relationship with the one of community biomass 
($r_\delta = `r tps_bm_sest["tlvl"]`$, Fig. \@ref(fig:bmr)B, Table
\@ref(tab:aov)), meaning
that stations where community biomass decreased over time, average trophic level
decreased over time and vice versa. It means that in communities losing biomass
over time, the average biomass of the food-web moved toward lower trophic level
and vice versa. The spatial gradient also showed a positive relationship between
median average trophic level and median community biomass, but twice less strong
than the temporal one ($r_\delta = `r sp_b_sest["tlvl"]`$, Fig. \@ref(fig:std),
Table \@ref(tab:aov-sp)). As for connectance, the temporal trends of average
trophic level was unrelated to the one of species richness. It means in
communities either losing or gaining species richness over time, the location of
average biomass along   trophic level did not significantly change.  However, we
found a strong negative relationship along the spatial gradient 
($r_\delta = `r sp_r_sest["tlvl"]`$, Fig. \@ref(fig:std), Table
\@ref(tab:aov-sp)).

Evenness of species biomass was negatively linked to community
biomass in time and along the spatial gradient 
($r_\delta = `r tps_bm_sest["p_bm"]`$ and $`r sp_b_sest["p_bm"]`$ resp.; Fig.
\@ref(fig:bmr)C, Fig. \@ref(fig:std)). In time, it means that food-webs that
lost biomass over time also had biomass more evenly distributed by species over
time and vice versa. A similar reasoning applies to spatial gradient. On the
other hand, we found a positive relationship between the temporal trends of
biomass evenness and the one of species richness (Fig. \@ref(fig:bmr)F, Table
\@ref(tab:aov), $r_\delta = `r tps_r_sest["p_bm"]`$). We found the same
relationship along the spatial gradient, twice strong however 
($r_\delta = `r sp_r_sest["p_bm"]`$).

@@

## BEF works in space and time

Spatially and temporally, species richness scaled positively with
community biomass. Spatially, this result is in accordance with numerous studies was
already found in streams [@woods_testing_2020; @danet_species_nodate] but also
in grasslands, salt marches and forests [@hector_plant_1999;
@tilman_biodiversity_2014; @wu_relationship_2015; @li_relationship_2018; but see
@grace_integrative_2016]. Therefore, our results confirm that BEF prediction
works also in time. The positive relationship of species richness on total
biomass might be due to resource use complementarity [@carey_determining_2011],
which is held by niche differentiation [@loreau_biodiversity_1998;
@loreau_partitioning_2001] or to the higher probability of getting high biomass
species with higher species richness [@loreau_biodiversity_2001].

@@

We found that the temporal increase of the average trophic level was positively
linked to increase of community biomass in line with previous studies suggesting
that the presence of higher predators can lead to higher community biomass
[@wang_biodiversity_2018; @danet_species_nodate]. Increase of average trophic
level can be associated to either gain of biomass in the highest trophic level
or losses of biomass in the lowest trophic. In our case, such gain or losses of
average trophic level is probably associated to the variations of the presence
of large fishes. The finding that the evenness of biomass, measured by the
pielou, decreased through time as community biomass increased and reversely,
meaning that biomass distribution across species become more even when community
biomass decreased and reversely that biomass distribution across species become
less even when community biomass increased. In the same line, the temporal
trends of community biomass were positively correlated to the temporal trends of
the proportion of piscivorous species and trophic species in the community (Fig.
SXX).

We also found in space that evenness of biomass was positively related to
community biomass while evenness of abundances had no relation (Fig. SXX). It
constrasts with previous studies which found a negative relationship between
biomass and evenness of abundances in marine and plant communities
[@drobner_relation_1998; @maureaud_biodiversityecosystem_2019] or studies that
found more even plant communities in terms of abundances number display higher
community biomass [@wilsey_biodiversity_2000].

We further found that the temporal trends of connectance was positively
linked to that one of community biomass, meaning that connectance increases through
time in a given site when community biomass increases through time and reversely.
This is in line with the fact that the fishes located at higher trophic level
also display a larger number of links as the size of the predation window
increases with the size of the fishes. This result is also in accordance with  
theoretical work showing that  higher food-web connectance can result in higher
community biomass [@thebault_food-web_2003]. But it contrasts with previous
results [@poisot_trophic_2013; @danet_species_nodate], highlighting that the
relation between connectance and community biomass may be dependant of assembly
processes at work.

@@

In contrasts to the results for the community biomass, the temporal trends of
evenness of biomass was positively related to the temporal trends of species
richness, meaning that as species richness temporally increase, the
distribution of biomass become more even, while temporal trends of evenness of
abondances had no significant relationship with species richness (Fig. SXX).
Those results seem indicate that biomass trends are linked to few big individuals
while richness trends are linked to overall shift in dominance in communities.

Surprinsingly, we found that the temporal trends of connectance was not related to
the one of species richness as found spatially in most empirical food-web
[@dunne_network_2006] and also in the present study. This result suggests that
the structure of food-webs might not be very sensitive to variation of species
richness and that species loss and gain concerns species of low trophic level,
which in turn affect connectance very little. Remarkably, temporal
variations of community biomass and the ones of species richness seem to be
linked to high trophic level in the first while being linked to lower trophic
level in the second. Those results highlight that variations in community
biomass and species richness are linked to different assembly processes and/or
pressures, but it is still an open question.

@@

In contrast with a meta-analysis [@soininen_relationship_2012], we found that
the evenness of abundances is positively related to the one of species
richness in space in stream fish assemblages but found no relationship in time.

This result has profound implications for conservation. As more and more
studies report community biomass and species richness trends [@hallmann_more_2017;
@ceballos_biological_2017], we documented that they might not be
equivalent in terms of consequences on the structure of the community. In
particular, the loss of the highest trophic level associated to the loss of
community biomass could of have dramatic consequences on ecosystem functioning
as highest predators are prone to stabilise communities
[@shanafelt_stability_2018; danet_species_nodate]. In contrasts, the loss of
species richness was poorly associated to modification of food-web structure,
suggesting that the species loss and gain concern lower trophic levels, which
are also in transmitting energy toward higher level of the trophic chain.

@@

```{r}
fig_mat_cap <- paste0(
  "(A) Distribution of the fishing sites. (B-F) Temporal trends of biomass,
  species richness, connectance, average trophic level and evenness of biomass
  for two stations. (À retravailler)"
)
```
```{r, fig.cap=fig_mat_cap, out.width="100%"}
knitr::include_graphics(get_mypath("figures", "mat_met_inkscape.pdf"))
```
